---
title: "sensor_ag"
author: "Denver Link"
date: "2026-06-16"
output: html_document
editor_options: 
  chunk_output_type: console
---

# library
```{r}
library(tidyverse)
```

# data
```{r}
#mn DNR
dnr_do <- read_csv(file.path("G:",
                         "Shared drives",
                         "Hansen Lab",
                         "RESEARCH PROJECTS",
                         "LCCMR Coldwater Habitat Watersheds - MM",
                         "Data",
                         "Raw data",
                         "Continuous temp DO logger data",
                         "Sentinel Lakes 2026-03-10",
                         "do_data.csv"))

dnr_temp <- read_csv(file.path("G:",
                         "Shared drives",
                         "Hansen Lab",
                         "RESEARCH PROJECTS",
                         "LCCMR Coldwater Habitat Watersheds - MM",
                         "Data",
                         "Raw data",
                         "Continuous temp DO logger data",
                         "Sentinel Lakes 2026-03-10",
                         "temp_data.csv"))

dnr_metadata = read_csv(file.path("G:",
                         "Shared drives",
                         "Hansen Lab",
                         "RESEARCH PROJECTS",
                         "LCCMR Coldwater Habitat Watersheds - MM",
                         "Data",
                         "Raw data",
                         "Continuous temp DO logger data",
                         "Sentinel Lakes 2026-03-10",
                         "logger_metadata.csv"))

#voya
temp_21 <- read_csv(file.path("G:",
                         "Shared drives",
                         "Hansen Lab",
                         "RESEARCH PROJECTS",
                         "LCCMR Coldwater Habitat Watersheds - MM",
                         "Data",
                         "Raw data",
                         "Continuous temp DO logger data",
                         "VOYA 2026-03-24",
                         "VOYA_temp_21.csv"))

temp_22 <- read_csv(file.path("G:",
                         "Shared drives",
                         "Hansen Lab",
                         "RESEARCH PROJECTS",
                         "LCCMR Coldwater Habitat Watersheds - MM",
                         "Data",
                         "Raw data",
                         "Continuous temp DO logger data",
                         "VOYA 2026-03-24",
                         "VOYA_temp_22.csv"))
```

# data ag dnr
```{r}
#grab the notes from the metadatas
dnr_metadata_notes <- dnr_metadata %>% 
  select(basin_id, 
         station,
         deployment,
         logger_model,
         notes)

#add the notes to the temp data
dnr_temp <- dnr_temp %>% 
  left_join(dnr_metadata_notes, by = c("basin_id",
                                       "station",
                                       "deployment",
                                       "logger_model"),
            relationship = "many-to-one")

dnr_temp_trial <- dnr_temp %>%
  #create units column
  #create a logger notes column as an aggregation of a couple of columns,
  #rename the metadata notes column to survey notes
  #rename columns for aggregation
  #location notes as deployment number
  mutate(depth_units = "m",
         temp_units = "c",
         logger_notes = paste0("serial_number: ",
                               serial_number,
                               ": logger_model: ",
                               logger_model),
         location_notes = paste0("deployment_number: ", deployment),
         original_file = paste0("temp_data.csv", " ;", "logger_metadata.csv")) %>% 
  select(lake_name = basin_name,
         lake_id = basin_id,
         station_id = station,
         date_time = standardized_date_time,
         depth = depth_m,
         depth_units,
         temp = temp_c,
         temp_units,
         logger_notes,
         survey_notes = notes,
         location_notes,
         flag_temp = flaggross,
         original_file)
rm(dnr_temp, dnr_metadata)

#do
dnr_do <- dnr_do %>% 
  left_join(dnr_metadata_notes, by = c("basin_id",
                                       "station",
                                       "deployment",
                                       "logger_model"),
            relationship = "many-to-one")

dnr_do_trial <- dnr_do %>%
  #create units column
  #create a logger notes column as an aggregation of a couple of columns,
  #rename the metadata notes column to survey notes
  #rename columns for aggregation
  #location notes as deployment number
  mutate(depth_units = "m",
         do_units = "mgl",
         temp_units = "c",
         logger_notes = paste0("serial_number: ",
                               serial_number,
                               ": logger_model: ",
                               logger_model),
         location_notes = paste0("deployment_number: ", deployment),
         original_file = paste0("do_data.csv", " ;", "logger_metadata.csv")) %>% 
  select(lake_name = basin_name,
         lake_id = basin_id,
         station_id = station,
         date_time = standardized_date_time,
         depth = depth_m,
         depth_units,
         do = do_mgl,
         do_units,
         temp = temp_c,
         temp_units,
         logger_notes,
         survey_notes = notes,
         location_notes,
         flag_do = flaggross_do,
         flag_temp = flaggross_temp,
         original_file)
rm(dnr_do, dnr_metadata_notes)

#join any rows of a common lake, station, depth, time for temp/do
check <- dnr_do_trial %>% 
  inner_join(dnr_temp_trial, by = c("lake_id", "station_id", "depth", "date_time"), suffix = c("_do", "_temp")) %>% 
  select(lake_id, station_id, depth, date_time, logger_notes_do, logger_notes_temp, location_notes_do, location_notes_temp, temp_do, temp_temp)

check  %>% 
  distinct(lake_id, station_id)
```

# adding some arrow friendly calculations for Mike
The goal here is to make some arrow friendly code that can visualize lake profiles (interpolated across depths) that has additional information displayed - tracking the depth of a certain temperature over time. 

In talking to Mike, he suggested that he interpolates between depth sensor with a linear line for every 0.1 m. A linear fit is probably fine because sensor have been positioned carefully to avoid missing a "curve", but that still can't be ruled out. We should consider a better fitting strategy.

To make this work with arrow, I have come up with a general workflow:
1. read in standardized parquet 
2. filter to lakes and time periods of interest 
   -this can be generally as large as your memory will allow, but it probably just makes sense to limit it to what you are plotting at the given moment - looking at a two year stretch in carols VS all lakes and years
3a. isotherm analysis 
     - direct interpolation between observations where the temperature of interest is "sandwiched" 
     - create isotherm depth over time (rows are date-time, column is "depth at temp X", value is depth)
3b. heat map data
    -temporary 0.1 meter depth interpolation (making sure we filter for place and time we are interested in)
    - vis
```{r}
library(tidyverse)
library(arrow)
library(lubridate)

#lets start by reading in he parquet files OR hived data
temp_all <- open_dataset(file.path("G:",
                                   "Shared Drives",
                                   "Hansen Lab",
                                   "LCCMR",
                                   "Data",
                                   "Path",
                                   "To",
                                   "File"))

# filter to lakes and times of interest 
temp_all %>% 
  filter(lake %in% c(LAKES),
         datetime >= START_DATE,
         datetime <= END_DATE
         ) %>% 
  select(
    lake,
    station,
    datetime,
    depth_m,
    temp_c
  ) %>% 
  collect()


#### function for finding depth of some temperatures 
temps_to_track <- c(10, 15, 20)

find_isotherms <- function(data, target_temps) {
  data <- data %>% 
    #make sure no NAs in depth or temp data
    filter(!is.na(depth_m),
           !is.na(temp_c)) %>% 
    #arrage data by depth
    arrange(depth_m)
  
  if (nrow(data) < 2) {
    return(tibble())
  }
  
  map_dfr(target_temps, function(target_temp) {
    
    profile <- data %>% 
      mutate(
        temp_diff = temp_c - target_temp,
        temp_diff_next = lead(temp_diff),
        depth_next = lead(depth_m),
        temp_next = lead(temp_c)
      ) %>% 
      filter(
        !is.na(temp_diff_next),
        temp_diff * temp_diff_next <= 0,
        temp_next != temp_c
      )
    
    if (nrow(profile) == 0) {
      return(tibble(
        target_temperature = target_temp,
        isotherm_depth = NA_real_
      ))
    }
    
    profile %>% 
    slice(1) %>% 
    transmute(
      target_temperature = target_temp,
      
      isotherm_depth = 
        depth_m +
        (target_temp - temp_c) /
        (temp_next - temp_c) *
        (depth_next - depth_m)
    )
  })
}

#test on fake data
test_profile <- tibble(
  lake = c(
    # Carlos - 12:00
    rep("carlos", 5),
    # Carlos - 12:10
    rep("carlos", 5),
    # Ten Mile - 12:00
    rep("tenmile", 6),
    # Ten Mile - 12:10
    rep("tenmile", 6)
  ),
  
  datetime = c(
    # Carlos - 12:00
    rep(as.POSIXct("2026-07-01 12:00:00"), 5),
    # Carlos - 12:10
    rep(as.POSIXct("2026-07-01 12:10:00"), 5),
    # Ten Mile - 12:00
    rep(as.POSIXct("2026-07-01 12:00:00"), 6),
    # Ten Mile - 12:10
    rep(as.POSIXct("2026-07-01 12:10:00"), 6)
  ),
  
  depth_m = c(
    # Carlos - 12:00
    1, 3, 5, 8, 11,
    
    # Carlos - 12:10
    1, 3, 5, 8, 11,
    
    # Ten Mile - 12:00
    1, 3, 6, 9, 11, 15,
    
    # Ten Mile - 12:10
    1, 3, 6, 9, 11, 15
  ),
  
  temp_c = c(
    # Carlos - 12:00
    22, 21, 18, 10, 7,
    
    # Carlos - 12:10
    22, 20, 17, 9, 6,
    
    # Ten Mile - 12:00
    23, 21, 18, 11, 7, 4,
    
    # Ten Mile - 12:10
    23, 21, 17, 10, 6, 3
  )
)

#run the function - we must group by lake and time
isotherms <- test_profile %>% 
  group_by(
    lake, 
    datetime
  ) %>% 
  group_modify( ~ find_isotherms(
    .x,
    target_temps = temps_to_track
  )) %>% 
  ungroup()

test_profile %>% 
  ggplot() +
  geom_point(aes(temp_c, depth_m)) +
  geom_line(aes(temp_c, depth_m, group = datetime)) +
  geom_smooth(aes(temp_c, depth_m, group = datetime)) +
  geom_point(data = isotherms, aes(target_temperature, isotherm_depth), color = "red") +
  scale_y_reverse() +
  facet_wrap(~lake)
# we can see here that we calculated the depth at a temperature on a linear line between two points with data from two different lakes that have a different number of sensors at different depths 

###########################################################################
#okay now lets do the heat map 
interpolate_profile <- function(data, depth_grid) {
  
  data <- data %>%
    filter(
      !is.na(depth_m),
      !is.na(temp_c)
    ) %>%
    arrange(depth_m)
  
  if (nrow(data) < 2) {
    return(tibble())
  }
  
  # Remove duplicate depths
  data <- data %>%
    distinct(depth_m, .keep_all = TRUE)
  
  # Linear interpolation
  # rule = 1 means don't extrapolate beyond observed depths
  approx(
    x = data$depth_m,
    y = data$temp_c,
    xout = depth_grid,
    method = "linear",
    rule = 1
  ) %>%
    as_tibble() %>%
    rename(
      depth_m = x,
      temp_c = y
    )
}

#set the depth grid to go from max depth to 0 by 0.1
# i am reusing the test data from above
heatmap_data <- test_profile %>%
  group_by(lake, datetime) %>%
  group_modify(
    ~ interpolate_profile(
        .x,
        depth_grid = seq(
          0,
          max(.x$depth_m, na.rm = TRUE),
          by = 0.1
        )
      )
  ) %>%
  ungroup()


#lets plot the heatmap and dots together
ggplot() +
  geom_raster(data = heatmap_data,
  aes(
    x = datetime,
    y = depth_m,
    fill = temp_c
  )) +
  
  geom_line(
    data = isotherms,
    aes(
      x = datetime,
      y = isotherm_depth,
      color = factor(target_temperature),
      group = interaction(lake, target_temperature)
    ),
    linewidth = 0.8
  ) +
  
  scale_y_reverse() +
  
  scale_fill_viridis_c(
    name = "Temperature (°C)",
    na.value = "white"
  ) +
  
  scale_color_manual(
    name = "Isotherm (°C)",
    values = c(
      "10" = "white",
      "15" = "black",
      "20" = "red"
    )
  ) +
  
  labs(
    x = "Date",
    y = "Depth (m)",
    title = "Lake temperature profile"
  ) +
  
  theme_bw() +
  facet_wrap(~lake)
```

I left off after creating the heat map. We will need to explore how much data we can pull in and still create the heat maps plots - we probably only viz one year at a time? that would be taken care of in the filtering prior to pulling data in